### 2023-04-12 research/v2.tex
Gitignore and resarch LaTex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{pmboxdraw}
\usepackage{minted}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\DeclareSymbolFont{letters}{OML}{ztmcm}{m}{it}

\begin{document}
\section{Introduction}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

Transcranial magnetic stimulation (TMS) is a common, non-invasive experimental technique used to evoke action potentials in cortical regions of the brain. In particular, researchers often target the motor cortex and measure the motor evoked potential (MEP) aroused by the stimulation. When the purpose of the experiment is to inquire on neuroplasticity, such stimulations are performed under a \textit{paired-pulse paradigm}.

The \textit{paired-pulse paradigm} (or \textit{double pulse paradigm}) consists in eliciting a series of two temporally proximate pulses (in the order of miliseconds). The evoked potentials of the double-stimulations are compared to those of single test stimulations, and their relative amplitude is taken as a proxy of neuroplasticity in the brain. The time separating each of the paired-pulses is termed \textit{interstimulous interval} (ISI). It is the general case that low intervals (4 or 5 miliseconds) produce intracortical inhibition, with the evoked potentials of paired stimulations being generally lower than those of single pulses. Greater intervals, on the other hand, tend to produce facilitation.

In the context of this paper, we shall term any coefficient that serves to represent the proportional relationship between the potentials evoked by paired and test pulses a \textit{measure of relative amplitude}. Measures of relative amplitude in neuroscience are generally computed at the subject level. This is reasonable, since hypotheses generally deal with differences across subject groups. However, TMS experimental results are pulse-specific evoked potentials, and transforming them to subject- or group-specific measures implies lowering the data resolution (i.e. representing the same phenomena with less information). In times when scientific inquiry can greatly benefit from forms of artificial intelligence that rely on large amounts of data, it is limiting to count only with low-resolution measures.

Hence, the goal of this paper is to provide pulse-specific measures of relative amplitude. We formulate coefficients that represent the relative amplitude of individual paired-pulses, keeping data resolution at its highest. Thus, from a computational perspective, our objective is constrained to the sphere of feature engineering. This allows for otherwise unfeasible computational analyses of TMS data, such as the use of machine learning models for the detection of different pulse-response patterns among different groups of clinical subjects.

We show such features allow for a machine learning classifier to correctly determine whether a subject belongs to one of four clinical categories based only on its evoked potentials across different inter-stimulous intervals with an accuracy of up to $90\%$.

@@

For simplicity, we will deal with the hypothetical situation in which a single ISI was used for paired stimulations. All of our results generalize to different ISI.

@@

Let $k$ be the number of experimental subjects in some subject group, to each of whom $n$ paired stimulations and $m$ test stimulations were elicited.

@@

    \begin{equation}
    \textbf{P} := \begin{bmatrix}x_{11} & x_{12}& \cdots & x_{1k} \\
    x_{21} & x_{22}& \cdots & x_{2k} \\
    & & ... & \\
    x_{n1} & x_{n2}& \cdots & x_{nk} 
    \end{bmatrix} ~~
    \textbf{T}:= \begin{bmatrix}t_{11} & t_{12}& ... & t_{1k} \\
    t_{21} & t_{22}& \cdots & t_{2k} \\
    & & ... & \\
    t_{m1} & t_{m2}& \cdots & t_{mk} 
    \end{bmatrix}
    \end{equation}
\end{definition}

\begin{definition}
Let $\mu_t := \begin{bmatrix} \mu_1, \mu_2, \ldots, \mu_k \end{bmatrix}^\top$ be a vector where $\mu_i$ is the average test potential for the $i$th subject. In other words,

@@

\begin{remark}
    Evoked potentials of both paired and test pulses conform to a Gamma distribution, as discussed in the appendix (?). This means it is always assumed $\forall x \in \textbf{P}, \forall t\in \textbf{T}|x, t \in \mathbb{R}^+$.
\end{remark}

@@

\begin{definition}
Let $x \in \textbf{P}_{\star i}$ be a single paired-stimulation MEP in the $i$th experimental subject, and $\textbf{t} = \textbf{T}_{\star i}$ the vector containing all test MEPs of that subject. Then we define two pulse-specific relative amplitude measures,

@@

\begin{remark}
    $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$. (For a proof of this property, consult the appendix.)
\end{remark}

Notice that $\rho$ is the proportion between the potential $x$, evoked by a paired stimulation, with respect to the average potential of single test stimulations. On the other hand, $\delta$ is the average proportion of $x$ with respect to each single test pulse.

The $\rho$ function is not an entirely new contribution. The traditional group level relative amplitude measure is conceived as the average, across all subjects in a group, of the average paired response divided by the average test response of each subject. If we let $S_i$ be the average relative amplitude of the $i$th subject, then $S_i$ has ben traditionally defined as

@@

as it is easy to see from decomposing the sum in the numerator. In other words, the traditionally used measure of relative amplitude, at the subject level, has always been the average $\rho$ in a subject.

@@

As stated earlier, action potentials evoked by trasncranial magnetic stimulation follow a Gamma distribution; one that closely resembles an exponential distribution (see \textbf{Statistics} section). A valid form of data augmentation that also attempts to deal with the presence of outliers is to weight the averages involved in $\rho, \delta$ using inverse-variance weights. The attempt is to allow for potentials that lay at the tail of the distribution to contribute in a way proportional to their probability.

@@

We used data collected across $N = ?$ subjects at the \textit{Laboratory for the Study of Slow-wave sleep}, University of Pennsylvania, with $H = ?$ healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD). Transcranial magnetic stimulation of the motor cortex was elicited to them after a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep. Motor evoked potentials were measured via an electrode (?) placed on the subjects' thumb. In the slow-wave disruption session, an auditory stimulus with sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet not strong enough to wake the subjects, was elicited. This experimental setting produces four distinct categories, two depending on the subject group and two on the type of sleep session underwent.

@@

The exact line of code that defined the model and its parameters was

\begin{minted}{python}
    bag_clf = AdaBoostClassifier(
        DecisionTreeClassifier(max_depth=8), 
        n_estimators=20, 
        algorithm='SAMME.R', 
        learning_rate=0.01)
\end{minted}

@@

\section{Appendix}

@@

\textbf{Theorem: } $\forall x \in \mathbb{R}^+|\delta(x) \geq \rho(x)$

@@

\begin{align} \delta(x) &= \frac{x}{m}\sum^m\frac{1}{t_j} \\ \rho(x) &= \frac{xm}{\sum^m t_j} \end{align}

@@

\begin{align} \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\
\iff S_2^mS_1^m &\geq m^2
\end{align}

@@

\begin{equation}
S^{k+1}_1 S^{k+1}_2 \geq (k+1)^2
\end{equation}

\begin{align} S^{k+1}_1 S^{k+1}_2 &\geq (k+1)^2 \\
(S_1^k+\frac{1}{t_{t+1}})(S_2^k+t_{k+1}) &\geq k^2+2k+1 \\
S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2+1 &\geq k^2+2k+1 \\
S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2 &\geq k^2+2k
\end{align}

@@

\begin{align}\frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}} &\geq 2k \\
\iff\frac{1}{t_{k+1}}(t_1+t_2+...+t_k)+t_{k+1}(\frac{1}{t_1}+\frac{1}{t_2}+...+\frac{1}{t_k}) &\geq 2k \\
\iff \Big(\frac{t_1}{t_{k+1}}+\frac{t_2}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+\frac{t_{k+1}}{t_2}+...+\frac{t_{k+1}}{t_k}\Big) &\geq 2k \\
\iff \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+... + n+\frac{1}{n}}^{\text{$2k$ terms} } &\geq 2k
\end{align}

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{pmboxdraw}
\usepackage{graphicx} \graphicspath{
{./images/} }
\usepackage{indentfirst}
\usepackage{multicol}
\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
\DeclareSymbolFont{letters}{OML}{ztmcm}{m}{it}
\newtheorem{theorem}{Theorem}

\begin{document}

\begin{multicols}{2}
\section{Introduction}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

Transcranial magnetic stimulation (TMS) is a common,
non-invasive experimental technique used to evoke action
potentials in cortical regions of the brain. In particular,
researchers often target the motor cortex and measure the
motor evoked potential (MEP) aroused by the stimulation.
When the purpose of the experiment is to inquire on
neuroplasticity, such stimulations are performed under the
\textit{paired pulse paradigm}.

The \textit{paired pulse paradigm} (or \textit{double pulse
paradigm}) consists in eliciting a series of two temporally
proximate pulses (in the order of miliseconds). The evoked
potentials of the double-stimulation are compared to those
of single test stimulations, and their relative amplitude is
taken as a proxy of neuroplasticity in the brain. The time
separating each of the paired pulses is termed
\textit{interstimulus interval} (ISI). It is the general
case that low intervals (4 or 5 miliseconds) produce
intracortical inhibition, with the evoked potentials of
paired stimulations being generally lower than those of
single pulses. Greater intervals, on the other hand, tend to
produce facilitation.

In the context of this paper, we shall term any coefficient
that serves to represent the proportional relationship
between the potentials evoked by paired and test pulses a
\textit{measure of relative amplitude}. Measures of relative
amplitude in neuroscience are generally computed at the
subject level. This is reasonable, since hypotheses
generally deal with differences across subject groups.
However, TMS experimental results are pulse-specific evoked
potentials, and transforming them to subject- or
group-specific measures implies down-scaling the data resolution.

The goal of this paper is to provide pulse-specific measures of relative
amplitude. This is, coefficients that represent the relative amplitude of each
individual paired pulse with respect to the set of test pulses in an
experimental session. The purpose of this endeavor is to keep data resolution
at its highest, which on its turn allows for the implementation of data-driven
artificial intelligence in the analysis of the 
experimental results. Thus, from a computer science perspective, our objective is
constrained to the sphere of feature engineering.

We will show how pulse-specific measures of relative amplitude allow for
otherwise unfeasible computational analyses of TMS data, such as the use of
machine learning models for the detection of different pulse-response patterns
among different groups of clinical subjects. In particular, we will show they
allow for a machine learning classifier to correctly determine whether a
subject belongs to one of four clinical categories, based only on its evoked
potentials and across different inter-stimulus intervals, with an accuracy of up to
$90\%$.

@@

For simplicity, we will deal with the hypothetical situation
in which a single ISI was used for paired stimulations. All
of our results generalize to different inter-stimulus intervals.

@@

Let $k$ be the number of experimental subjects in some
subject group $\mathcal{G}$, to each of whom $n$ paired stimulations and
$m$ test stimulations were elicited.

@@

\begin{equation*} \textbf{P} := \begin{bmatrix}x_{11} &
    x_{12}& \ldots & x_{1k} \\ x_{21} & x_{22}& \ldots &
    x_{2k} \\ \vdots & \vdots &\ddots& \vdots \\ x_{n1} & x_{n2}& \cdots & x_{nk}
\end{bmatrix} ~~ \textbf{T}:= \begin{bmatrix}t_{11} &
t_{12}& \ldots & t_{1k} \\ t_{21} & t_{22}& \cdots & t_{2k} \\
      \vdots & \vdots &\ddots& \vdots \\ t_{m1} & t_{m2}& \cdots & t_{mk}
\end{bmatrix} \end{equation*} \end{definition}

\begin{definition} 
    Let $\mu_t := \begin{bmatrix} \mu_1 & \mu_2 & \ldots & \mu_k
    \end{bmatrix}^\top$ be such that $\mu_i$ is the average test potential for
    the $i$th subject. In other words,

@@

\begin{remark} 
    Evoked potentials of both paired and test pulses conform to a Gamma
    distribution, as discussed in the appendix (?). This means it is always an
    implicit assumption that $\langle\forall x, t : x \in \textbf{P}, t\in \textbf{T}
    : x, t \in \mathbb{R}^+\rangle$. 
\end{remark}

@@

\begin{definition} 
    Let $x \in \textbf{P}_{\star i}$ be the MEP of a single paired stimulation
    elicited on the $i$th experimental subject, and $\textbf{t} =
    \textbf{T}_{\star i}$ the vector containing the MEPs of all test pulses
    elicited on that subject. Then we define two pulse-specific relative
    amplitude measures,

@@

\begin{remark} 
    $\forall x: x \in \mathbb{R}^+:\delta(x) \geq \rho(x)$. (For a proof of this
    property, consult the appendix.) 
\end{remark}

Notice that $\rho(x)$ is the proportion between the potential
$x$, evoked by a paired stimulation, with respect to the
average potential of single test stimulations. On the other
hand, $\delta(x)$ is the average proportion of $x$ with respect
to each single test pulse.

The $\rho$ function is not an entirely new contribution. Traditionally, the
group-level relative amplitude measure is conceived as the average, across all
subjects in a group, of the average paired response divided by the average test
response. This means that, if we let $S_i$ be the average relative amplitude of
the $i$th subject, then $S_i$ has been traditionally defined as

@@

as it is easy to see from decomposing the sum in the
numerator. In other words, the traditionally used measure of
relative amplitude, at the subject level, has always been
the average $\rho$ in a subject, albeit it was never defined explicitly.

@@

As stated earlier, action potentials evoked by transcranial
magnetic stimulations follow a Gamma distribution; one that
closely resembles an exponential distribution (see
\textbf{Statistics} section). A valid form of data
augmentation that also attempts to deal with the presence of
outliers is to weight the averages involved in $\rho,
\delta$ using inverse-variance weights. The attempt is to
allow for potentials that lay at the tail of the
distribution to contribute in a way proportional to their
probability.

@@

We used data collected across $N = ?$ subjects at the \textit{Laboratory for the
Study of Slow-wave sleep activity}, University of Pennsylvania, with $H = ?$
healthy controls and $D = ?$ diagnosed with major depressive disorder (MDD).
Transcranial magnetic stimulation of the motor cortex was elicited to them after
a night of baseline sleep and after a night of slow-wave disruption (SWD) sleep.
Motor evoked potentials were measured via an electrode (?) placed on the
subjects' thumb. In the slow-wave disruption session, an auditory stimulus with
sufficient strength to interrupt the normal occurrence of slow-wave sleep, yet
not strong enough to wake the subjects, was elicited. This experimental setting
produces four distinct categories, two depending on the subject group and two on
the type of sleep session underwent.

@@

@@

\pagebreak
\section{Appendix}

@@

\begin{theorem}
    $\forall x: x \in \mathbb{R}^+ : \delta(x) \geq \rho(x)$
\end{theorem}

@@

\begin{align*} \delta(x) &= \frac{x}{m}\sum^m\frac{1}{t_j} \\
                \rho(x) &= \frac{xm}{\sum^m t_j} \end{align*}

@@

\begin{align*} 
    \frac{x}{m}\sum^m\frac{1}{t_j} &\geq \frac{xm}{\sum^m t_j} \\ 
    S_2^mS_1^m &\geq m^2
\end{align*}

@@

\begin{equation*} 
    S^{k+1}_1 S^{k+1}_2 \geq (k+1)^2
\end{equation*}

This can be proven as follows.

\begin{align*} 
    S^{k+1}_1 S^{k+1}_2 &\geq (k+1)^2 \\ 
    (S_1^k+\frac{1}{t_{t+1}})(S_2^k+t_{k+1}) &\geq k^2+2k+1 \\
    S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2+1 &\geq k^2+2k+1 \\
    S^k_1S^k_2+ t_{k+1}S_1^k + \frac{1}{t_{k+1}}S^k_2 &\geq k^2+2k 
\end{align*}

@@

\begin{align*}
    \frac{1}{t_{k+1}}\sum_{j=1}^mt_j+t_{k+1}\sum_{j=1}^m\frac{1}{t_{j}} &\geq 2k \\
    \Big(\frac{t_1}{t_{k+1}}+...+\frac{t_k}{t_{k+1}}\Big)+\Big(\frac{t_{k+1}}{t_1}+...+\frac{t_{k+1}}{t_k}\Big)
    &\geq 2k \\ \iff
    \overbrace{a+\frac{1}{a}+b+\frac{1}{b}+... +
    n+\frac{1}{n}}^{\text{$2k$ terms} } &\geq 2k
\end{align*}
